---
TITLE:  Reproducible research - Assignment 1
AUTHOR: Sai Prasad Guttapalli
DATE: 12/21/2018
output: html_document
keep_md: true
---

## Loading and preprocessing the data
Read the data file 'activity.csv'
convert date from factor format to date format 'ymd
```{r, echo=TRUE} 
library(lubridate)
library(dplyr)
library(ggplot2)
actv <- read.csv('~/desktop/activity.csv')        
actv$date <- ymd(as.Date(as.character(actv$date)))       
a<- actv          
actv <- filter(actv, !is.na(steps) )  
```  

## What is mean total number of steps taken per day?
Calculate the total number of steps taken per day
Make a histogram of the total number of steps taken each day
Calculate and report the mean and median of the total number of 
steps taken per day

```{r, echo=TRUE}    
dailysteps <- actv %>% group_by(date) %>%   
              summarise(totalsteps = sum(steps))    
hist(dailysteps$totalsteps, freq = NULL, axes = T,     
           main ='steps taken every day', xlab =' Daily steps')    
mediandailysteps <- summary(dailysteps$totalsteps)[3]    
meandailysteps <- summary(dailysteps$totalsteps)[4]    
print(c('Mean Daily Steps =', meandailysteps))    
print(c('Median Daily Steps =', mediandailysteps))    
```  


## What is the average daily activity pattern?
Make a time series plot ("type="l") of the 5-minute interval (x-axis) 
and the average number of steps taken, averaged across all days (y-axis)
Which 5-minute interval, on average across all the days in the dataset, 
contains the maximum number of steps?

```{r, echo=TRUE}  
stp_int <- actv %>% group_by(interval) %>%      
                           summarise(avg_steps =  mean(steps) )    
  
plot(stp_int$interval, stp_int$avg_steps,     
     type = 'l', xlab = 'intervals', ylab = 'Average Steps',     
     main = 'Average steps over the day')    
    
mint <- stp_int %>%     
          filter(avg_steps == max(avg_steps) )    
print(c("The Interval in which Max steps are taken = ", mint[1]) )    
```  


## Imputing missing values
Note that there are a number of days/intervals where there are missing 
values (coded as NA).
Calculate and report the total number of missing values (i.e. NAs)

```{r, echo=TRUE}    
actv <- a    
missingvals <- sum(is.na(actv$steps))    
print(c("Total Missing Values", missingvals))    
  
for (i in 1:dim(actv)[1]) {    
     if (is.na(actv[i,1] ) )    
     {    
       const <- filter(stp_int, interval == actv[i,3])    
       actv[i,1] <- const[1,2]    
      }     
}    
  
dailysteps <- actv %>% group_by(date) %>% summarise(sum(steps))    
names(dailysteps)[2] <- 'totalsteps'    
hist(dailysteps$totalsteps, freq = NULL, axes = T,     
     main ='steps taken every day', xlab =' Daily steps')    
mediandailysteps <- summary(dailysteps$totalsteps)[3]    
meandailysteps <- summary(dailysteps$totalsteps)[4]    
print(c('Mean Daily Steps after replacement=', meandailysteps))    
print(c('Median Daily Steps after replacement =', mediandailysteps))   
  
```  


## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” 
and “weekend” indicating whether a given date is a weekday or weekend 
day. Make a panel plot containing a time series plot  (i.e.type="l") 
of the 5-minute interval (x-axis) and the average number of steps taken, 
averaged across all weekday days or weekend days (y-axis). 

```{r, echo=TRUE}   
actv <- mutate(actv, dday = case_when(     
    weekdays(date) %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "weekday",     
    weekdays(date) %in% c("Saturday", "Sunday") ~ "weekend" ) )    
  
stp_week <- actv %>% group_by(dday, interval) %>%     
            summarise(week_steps =  mean(steps) )    
par(mfrow = c(1,2) )    
qplot(interval, week_steps, data=stp_week, facets=.~dday,     
      geom = c("line"))  
```  
